Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-08 trunk/doc/thesis/report.tex
ARABIC TITLE GRRRRRR

%\chapter{ملخص}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides power to households, companies and organizations. Therefore the whole world's economy counts on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in the Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously, without the guidance of humans, to performs their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ration was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy usage. This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e. the Internet.

@@

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's side or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a demand for the two-way communication, which is to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This use of different technologies comes with a problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it. To do so, \gls{6lowpan} had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network and ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}, the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
This study aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other even though the problems seem separate, one cannot solve  the a problem without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks only support \gls{ipv4} and therefore the communication is not possible, and this was due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in the Chapter 4 from its architectural aspect of the component, and in Chapter 5 from its deployment point of view. 
\paragraph{}
It is to be noted that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis were not delivered on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption and this information is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system and this is done by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

\section{Outline of the Thesis}
The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context. This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
\paragraph{Chapter 3}
It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet and the future of Internet.
\paragraph{Chapter 4}
It will present the architecture of the test-bed. It will discuss the system from different aspects, and how were the problems stated in section 1.3 solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter will mostly discuss the system deployment. It will start by presenting the hardware used in this project, then move to the technology enablers. Afterward, Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
\paragraph{Chapter 6}
It will start first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered will be shown, analyzed and interpreted.
\paragraph{Chapter 7, 8}
These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be ameliorated and how it can be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old and its problems are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it as to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to new technology: the \glsreset{sg}\gls{sg}. Although it is still a research project and has not reached maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problem existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer, but households can now produce and even sell the extra electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make this part of it. Consequently, research has taken the direction to integrate the \gls{sg} within the \gls{iot} as most of its aspects favor  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} will be divided into three layers:

@@

Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}. Still, Zigbee has the exclusive advantage of comsuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices. 

@@

\begin{itemize}
\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
\item Mote sensor data collection TCP client
\item Mote appliance control TCP server
\item \gls{ipv4}/\gls{ipv6} Gatewaying Process
\item \gls{ipv6} over USB tunnel
\item mote \gls{ipv6} backstation
\item Sensor data storage process
\item HTTP sensor data sender
\item Mobile client
\end{itemize}

\section{General Architecture of the System}
\paragraph{}
The general architecture is depicted in figure \ref{fig:gen_architecture}. This figure shows the existing components and their interactions. Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
\begin{figure}[htbp]
\centering
\includegraphics[width=17cm,height=17cm,keepaspectratio]{img/general_architecture.jpg}
\caption{General architecture depicting the main components of the system}
\label{fig:gen_architecture}
\end{figure}
\paragraph{}
Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside. As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.40]{img/layered_view.jpg}
\caption{Layered view showing where every component appears in the \gls{osi} model}
\label{fig:layered_view}
\end{figure}

@@

\paragraph{}
Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed. In other words, what are the software components that exist on which hardware machine. This is the goal of the matrix apearing in table \ref{table:soft_hard_deployment}. It shows where very software component is deployed. From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.

\begin{table}
    \begin{tabular}{lllll}
    		  \hline
    Software component - Hardware component & Mote & Mote sink & Gateway & Mobile \\ \hline
    Mote sensor data collection TCP client  & X    & X         & ~       & ~      \\ 
    Mote appliance control TCP server       & X    & X         & ~       & ~      \\
    \gls{ipv4}/\gls{ipv6} Gatewaying Process   & ~    & ~         & X       & ~      \\
    IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
    Mote IPv6 backstation                   & ~    & X         & ~       & ~      \\
    Sensor data storage process             & ~    & ~         & X       & ~      \\
    HTTP sensor data sender                 & ~    & ~         & X       & ~      \\
    Mobile client                           & ~    & ~         & ~       & X      \\
    \end{tabular}
    \caption{Software-hardware deployment table}
    \label{table:soft_hard_deployment}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.8]{img/ct.jpg}
\caption{Current transformer}
\label{fig:current_transformer}
\end{figure}

\begin{table}
    \begin{tabular}{ll}
    \hline
    slope    & 66.7 \gls{a} \\ \hline
    Constant & 0 \gls{a}    \\ \hline
    \end{tabular}
    \caption{Current Transformer values}
        \label{table:CT_vals}
\end{table}
To transform the initial voltage into the current intensity, we use constant variables provided by the \gls{ct} manufacturer. But these value cannot be taken for granted. To measure them, we conduct an experiment to using the Voltmeter and Amperemeter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function. This function resulted in the following values: 
\begin{itemize}
\item $s = 66.70$
\item $u_0 = 0$
\end{itemize}
Therefore, the first transformation function is: \\
\boldmath{$I = 66.7U_0$} \\
I: Intensity in \glsreset{a}\gls{a} \\
$U_0$: The voltage read from the \gls{ct}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.35]{img/appliance_control.jpg}
\caption{TCP Server controling appliance}
\label{fig:appliance_control}
\end{figure}

\paragraph{}
The mote on itself does not have the necessary hardware capabilities to control an appliance. It needs to be equipped with a Data Acquisition board also known as Extension Board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the appliance. The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}. This board has PINs which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance. However, the power supply of the appliances from the mote is not straightforward.
\paragraph{}
To make a mote control the appliance, the mote will govern the supply of electricity by controlling a small circuit that is responsible for pipelining the power flowing to the appliance. In short, the mote is equipped with an extension board to which an electric circuit is attached and whose responsible for the flow of power going to the appliance. Thus, the turn Off command will open the circuit and the appliance will stop getting fed electricity.

@@

\section{Mote IPv6 Backstation}
\paragraph{} 
This component is deployed only the sink of the \glsreset{wsn}\gls{wsn}. This mote is able to communicate with the sensor nodes using Zigbee and with the gateway using \gls{usb}. The software component deployed in this sink is a packet forwarder. In fact, once it receives an \gls{ipv6} packet that is going outside the \gls{wsn}, it is forwarded through \gls{usb}.
\paragraph{}
The component's goal is to forward packets in and out of the \gls{wsn} while other sensor node do not need to worry about how the packet will be communicated. This mote as part of the components aiming to solve the heterogeneity problem, provides communication transparency.

@@

\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
The \gls{ipv4}/\gls{ipv6} Gatewaying Process is one of the components that hides the heterogeneity that exist in the system. This component mainly deals with transforming packets since it extracts the datagram from the packet and create a completely new one that holds the datagram originating from the first one and communicates it. This component is composed of three subcomponents:
\begin{itemize}
\item Packet sniffer
\item Packet transformer
\item Address translator
\end{itemize}
\paragraph{}
Figure \ref{fig:gateway_transformation} depicts the architecture of this subsystem by showing the different subcomponents and their interactions. Moreover, it illustrates the data flow within this subsystem by showing the inputs and outputs of the subsystem.

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{img/mts400ca.jpg}
\caption{Crossbow MTS400CA Sensor Board}
\label{fig:mts400}
\end{figure}
\paragraph{}
The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}. Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment. This programming language is event-oriented and its syntax is similar to C. In addition, C libraries can be used in \gls{nesc} programs.

@@

\subsection{Mobile Client}
\paragraph{}
The mobile used is a smart phone runnning Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.

@@

\section{Mote IPv6 Backstation}

>>>>

%\chapter{RL{ملخص}}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides power to households, companies and organizations. Therefore the whole world's economy depends on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in the Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use. This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e. the Internet.

@@

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This use of different technologies comes with a problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component. This will make of this additional component a middleman and that would make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
This study aims at solving the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

\section{Outline of the Thesis}
\paragraph{}
The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context. This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ. Finally, its presents the commonalities ad differences between this thesis and the work previously done.
\paragraph{Chapter 3}
This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
\paragraph{Chapter 4}
It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter mostly discusses the system deployment. It starts by presenting the hardware used in this project, then moves to the technology enablers. Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
\paragraph{Chapter 6}
It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
\paragraph{Chapter 7, 8}
These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}. Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer. Households can now produce and sell the surplus of electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it. Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} is divided into three layers:

@@

Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}. Still, Zigbee has the exclusive advantage of consuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices. 

@@

\begin{itemize}
\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
\item Mote sensor data collection TCP client
\item Mote appliance control TCP server
\item \gls{ipv4}/\gls{ipv6} Gatewaying Process
\item \gls{ipv6} over USB tunnel
\item mote \gls{ipv6} back station
\item Sensor data storage process
\item HTTP sensor data sender
\item Mobile client
\end{itemize}

\section{General Architecture of the System}
\paragraph{}
The general architecture is depicted in figure \ref{fig:gen_architecture}. This figure shows the existing components and their interactions. Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
\begin{figure}[htbp]
\centering
\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/general_architecture.jpg}
\caption{General architecture depicting the main components of the system}
\label{fig:gen_architecture}
\end{figure}
%\clearpage
\paragraph{}
Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside. As an answer to such a question, figure \ref{fig:layered_view} is here to show in what layer does every component operates.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.40]{img/layered_view.jpg}
\caption{Layered view showing where every component operates in the \gls{osi} model}
\label{fig:layered_view}
\end{figure}

@@

\paragraph{}
Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed. In other words, what are the software components that exist on which hardware machine. This is the goal of the matrix appearing in table \ref{table:soft_hard_deployment}. It shows where very software component is deployed. From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.

\begin{table}
    \begin{tabular}{lllll}
    		  \hline
    Software component - Hardware component & Mote & Mote sink & Gateway & Mobile \\ \hline
    Mote sensor data collection TCP client  & X    & X         & ~       & ~      \\ 
    Mote appliance control TCP server       & X    & X         & ~       & ~      \\
    \gls{ipv4}/\gls{ipv6} Gatewaying Process   & ~    & ~         & X       & ~      \\
    IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
    Mote IPv6 back station                   & ~    & X         & ~       & ~      \\
    Sensor data storage process             & ~    & ~         & X       & ~      \\
    HTTP sensor data sender                 & ~    & ~         & X       & ~      \\
    Mobile client                           & ~    & ~         & ~       & X      \\
    \end{tabular}
    \caption{Software-hardware deployment table}
    \label{table:soft_hard_deployment}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.7]{img/ct.jpg}
\caption{Current transformer}
\label{fig:current_transformer}
\end{figure}

\begin{table}
    \begin{tabular}{ll}
    \hline
    slope    & 66.7 \gls{a} \\ \hline
    Constant & 0 \gls{a}    \\ \hline
    \end{tabular}
    \caption{Current Transformer values}
        \label{table:CT_vals}
\end{table}
To transform the initial voltage into the current intensity, we use constant variables provided by the \gls{ct} manufacturer. But these value cannot be taken for granted. To measure them, we conduct an experiment to using the Voltmeter and Ampere meter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function. This function resulted in the following values: 
\begin{itemize}
\item $s = 66.70$
\item $u_0 = 0$
\end{itemize}
Therefore, the first transformation function is: \\
\boldmath{$I = 66.7U_0$} \\
I: Intensity in \glsreset{a}\gls{a} \\
$U_0$: The voltage read from the \gls{ct}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.35]{img/appliance_control.jpg}
\caption{TCP Server controlling appliance}
\label{fig:appliance_control}
\end{figure}

\paragraph{}
The mote on itself does not have the necessary hardware capabilities to control an appliance. It needs to be equipped with a Data Acquisition board also known as Extension Board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the appliance. The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}. This board has Pins which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance. However, the power supply of the appliances from the mote is not straightforward.
\paragraph{}
To make a mote control the appliance, the mote will govern the supply of electricity by controlling a small circuit that is responsible for pipelining the power flowing to the appliance. In short, the mote is equipped with an extension board to which an electric circuit is attached and whose responsible for the flow of power going to the appliance. Thus, the turn Off command will open the circuit and the appliance will stop getting fed electricity.

@@

\section{Mote IPv6 Back station}
\paragraph{} 
This component is deployed only the sink of the \glsreset{wsn}\gls{wsn}. This mote is able to communicate with the sensor nodes using Zigbee and with the gateway using \gls{usb}. The software component deployed in this sink is a packet forwarder. In fact, once it receives an \gls{ipv6} packet that is going outside the \gls{wsn}, it is forwarded through \gls{usb}.
\paragraph{}
The component's goal is to forward packets in and out of the \gls{wsn} while other sensor node do not need to worry about how the packet will be communicated. This mote as part of the components aiming to solve the heterogeneity problem, provides communication transparency.

@@

\section{IPv4/IPv6 Gatewaying Process}
The \gls{ipv4}/\gls{ipv6} Gatewaying Process is one of the components that hides the heterogeneity that exist in the system. This component mainly deals with transforming packets since it extracts the datagram from the packet and create a completely new one that holds the datagram originating from the first one and communicates it. This component is composed of three subcomponents:
\begin{itemize}
\item Packet sniffer
\item Packet transformer
\item Address translator
\end{itemize}
\paragraph{}
Figure \ref{fig:gateway_transformation} depicts the architecture of this subsystem by showing the different subcomponents and their interactions. Moreover, it illustrates the data flow within this subsystem by showing the inputs and outputs of the subsystem.

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{img/mts400ca.jpg}
\caption{Crossbow MTS400CA Sensor Board}
\label{fig:mts400}
\end{figure}
\paragraph{}
The WSN is run using \gls{tinyos} which is one of the most used open source \gls{os} in \glspl{wsn}. Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment. This programming language is event-oriented and its syntax is similar to C. In addition, C libraries can be used in \gls{nesc} programs.

@@

\subsection{Mobile Client}
\paragraph{}
The mobile used is a smart phone running Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.

@@

\section{Mote IPv6 Back station}
